---
title: "panel_model"
author: "Andrew Walters"
date: "6/23/2019"
output: pdf_document
---

## Libraries
```{r message=FALSE, warning=FALSE}
# visual
library(knitr)
library(kableExtra)
library(ggplot2)
library(ggpubr)
library(stargazer)
# data manipulation
library(dplyr)
library(reshape)
library(MASS)
# modeling
library(stats)
library(matrixStats)
library(plm)
library(lmtest)
library(summarytools)
```

## Load Data
```{r message=FALSE, warning=FALSE}
# clear workspace
rm(list = ls())

data_cleanup <- function(df, start_date, end_date) {
  df$date <- as.Date(df$date, format = "%Y-%m-%d")
  df <- subset(df, date >= start_date )
  df <- subset(df, date <=  end_date)
  return(df)
}

top_candidates <- function(df, n_candidates) {
  avg_polls <- df %>%
    group_by(candidate) %>%
    summarize(mean = mean(polling_percentage_normalized, na.rm = TRUE))
  avg_polls <- avg_polls[order(-(avg_polls$mean)),] 
  top_list <- avg_polls[1:n_candidates,]
  top_df <- subset(df, candidate %in% top_list$candidate)
  bottom_list <- avg_polls[n_candidates+1:nrow(avg_polls),]
  bottom_df <- subset(df, candidate %in% bottom_list$candidate)
  return_dfs <- list(top=top_df, bottom=bottom_df)
  return(return_dfs)
}

dem20 <- read.csv("../data/dem20_dataset.csv")
dem20 <- data_cleanup(dem20, as.Date("2019-01-01"), as.Date("2019-06-30"))
dem20_split <- top_candidates(dem20, 7)

dem20_weekly <- read.csv("../data/dem20_weekly_dataset.csv")
dem20_weekly <- data_cleanup(dem20_weekly, as.Date("2019-01-01"), as.Date("2019-06-30"))

rep16 <- read.csv("../data/rep16_dataset.csv")
rep16 <- data_cleanup(rep16, as.Date("2015-01-01"), as.Date("2016-12-31"))
rep16_split <- top_candidates(rep16, 5)

head(dem20)
head(dem20_split$top)
```

## Including Plots

```{r}
print("bottom")
unique(dem20_split$bottom$candidate)
print("top")
unique(dem20_split$top$candidate)
```

You can also embed plots, for example:

```{r pressure, echo=FALSE}
make_ggplots <- function(df, cycle, party, description) {
  
  legend_nrow <- ceiling(length(df$candidate)/3)
  
  gg_poll <- ggplot(df, aes(x=date, y=polling_percentage_normalized)) +
    geom_line(aes(colour=candidate)) +
    ggtitle(sprintf("%s %s %s Polling", cycle, description, party)) +
    xlab("Date") +
    ylab("Normalized Polls") +
    guides(fill=guide_legend(ncol = 2))
  
  gg_dcount <- ggplot(df, aes(x=date, y=individual_donations_normalized)) +
    geom_line(aes(colour=candidate)) +
    ggtitle(sprintf("%s %s %s Donnation Count", cycle, description, party)) +
    xlab("Date") +
    ylab("Normalized Donnation Count")
  
  gg_dsum <- ggplot(df, aes(x=date, y=individual_donation_amount_normalized)) +
    geom_line(aes(colour=candidate)) +
    ggtitle(sprintf("%s %s %s Donnation Sum", cycle, description, party)) +
    xlab("Date") +
    ylab("Normalized Donnation Sum")
  
  gg_sdcount <- ggplot(df, aes(x=date, y=small_donor_count_normalized)) +
    geom_line(aes(colour=candidate)) +
    ggtitle(sprintf("%s %s %s Small Donnation Count (less than $200)", cycle, description, party)) +
    xlab("Date") +
    ylab("Normalized Small Donnation Count")
  
  gg_tv <- ggplot(df, aes(x=date, y=tv_mentions_normalized)) +
    geom_line(aes(colour=candidate)) +
    ggtitle(sprintf("%s %s %s TV Mention Count", cycle, description, party)) +
    xlab("Date") +
    ylab("Normalized Cable TV Mentions")
  
  return_plots <- list(poll=gg_poll, dcount=gg_dcount, dsum=gg_dsum, sdcount=gg_sdcount, tv=gg_tv)
  return(return_plots)
}


# Show a time series of the speed limit laws
#speedLimitMelted <- melt(speedLimit, id=c("year"))
dem20_ggtop <- make_ggplots(dem20_split$top, "2020", "Democrat", "Top 7")
dem20_ggbottom <- make_ggplots(dem20_split$bottom, "2020", "Democrat", "Bottom")

# display plots
ggarrange(dem20_ggtop$poll, dem20_ggbottom$poll, ncol=1, nrow=2)
ggarrange(dem20_ggtop$dcount, dem20_ggbottom$dcount, ncol=1, nrow=2)
ggarrange(dem20_ggtop$dsum, dem20_ggbottom$dsum, ncol=1, nrow=2)
ggarrange(dem20_ggtop$sdcount, dem20_ggbottom$sdcount, ncol=1, nrow=2)
ggarrange(dem20_ggtop$tv, dem20_ggbottom$tv, ncol=1, nrow=2)
```

```{r}
formula_agg <- ""
ii = 0
for (col in colnames(dem20)) {
  formula_agg <- paste(formula_agg, col, sep="+")
  if(ii%%4==0) {
    formula_agg <- paste(formula_agg, "\n")
  }
  ii <- ii+1
}
print(formula_agg)
```



Panel Model: all candidates, weekly aggregate, nlp featrues on polls
```{r}
dem20_weekly.panel <- pdata.frame(dem20_weekly, c("candidate", "date"))
dem20_weekly.panel <- dem20_weekly.panel[!duplicated(rownames(dem20_weekly.panel)),]
dem20.nlp.fit <- plm(polling_percentage_normalized ~ 
  X2018.congress+X2020.democractic.party+X2020.democractic.primary+X2020.democratic.primary+X2020.election+aoc+assange+biden.busing+booker+border+
  busing+buttigieg+civil.rights+climate.change+congress+de.blasio+drugs+economic.foreign.policy+electoral.college+female.congresswomen+first.debate+gun.laws+
  harris+healthcare+hickenlooper+impeachment+inslee+iowa.caucus+iran+iran.afghanistan+israel+kavanaugh+lgbtq+marijuana+mccain+mcconnell+
  mueller.report+north.korea+o.rourke+prison.reform+puerto.rico+russian.interference+sanders+student.debt+supreme.court+tax.return+
  tech+trump+trump.racist+unions+us.economy..socialism.vs.captialism.+venezuela+warren+williamson+yang,
  data=dem20_weekly.panel, model="within")

summary(dem20.nlp.fit)
```

Panel Model: all candidates, nlp featrues on polls
```{r}
dem20.panel <- pdata.frame(dem20, c("candidate", "date"))
dem20.panel <- dem20.panel[!duplicated(rownames(dem20.panel)),]
dem20.nlp.fit <- plm(polling_percentage_normalized ~ 
  X2018.congress+X2020.democractic.party+X2020.democractic.primary+X2020.democratic.primary+X2020.election+aoc+assange+biden.busing+booker+border+
  busing+buttigieg+civil.rights+climate.change+congress+de.blasio+drugs+economic.foreign.policy+electoral.college+female.congresswomen+first.debate+gun.laws+
  harris+healthcare+hickenlooper+impeachment+inslee+iowa.caucus+iran+iran.afghanistan+israel+kavanaugh+lgbtq+marijuana+mccain+mcconnell+
  mueller.report+north.korea+o.rourke+prison.reform+puerto.rico+russian.interference+sanders+student.debt+supreme.court+tax.return+
  tech+trump+trump.racist+unions+us.economy..socialism.vs.captialism.+venezuela+warren+williamson+yang,
  data=dem20.panel, model="within")

summary(dem20.nlp.fit)
```

Panel Model: all candidates, nlp featrues on donnation count
```{r}
dem20.nlp.fit <- plm(individual_donations_normalized ~ 
  X2018.congress+X2020.democractic.party+X2020.democractic.primary+X2020.democratic.primary+X2020.election+aoc+assange+biden.busing+booker+border+
  busing+buttigieg+civil.rights+climate.change+congress+de.blasio+drugs+economic.foreign.policy+electoral.college+female.congresswomen+first.debate+gun.laws+
  harris+healthcare+hickenlooper+impeachment+inslee+iowa.caucus+iran+iran.afghanistan+israel+kavanaugh+lgbtq+marijuana+mccain+mcconnell+
  mueller.report+north.korea+o.rourke+prison.reform+puerto.rico+russian.interference+sanders+student.debt+supreme.court+tax.return+
  tech+trump+trump.racist+unions+us.economy..socialism.vs.captialism.+venezuela+warren+williamson+yang,
  data=dem20.panel, model="within")

summary(dem20.nlp.fit)
```


Panel Model: all candidates, nlp featrues on tv mentions
```{r}
dem20.nlp.fit <- plm(tv_mentions_normalized ~ 
  X2018.congress+X2020.democractic.party+X2020.democractic.primary+X2020.democratic.primary+X2020.election+aoc+assange+biden.busing+booker+border+
  busing+buttigieg+civil.rights+climate.change+congress+de.blasio+drugs+economic.foreign.policy+electoral.college+female.congresswomen+first.debate+gun.laws+
  harris+healthcare+hickenlooper+impeachment+inslee+iowa.caucus+iran+iran.afghanistan+israel+kavanaugh+lgbtq+marijuana+mccain+mcconnell+
  mueller.report+north.korea+o.rourke+prison.reform+puerto.rico+russian.interference+sanders+student.debt+supreme.court+tax.return+
  tech+trump+trump.racist+unions+us.economy..socialism.vs.captialism.+venezuela+warren+williamson+yang,
  data=dem20.panel, model="within")

summary(dem20.nlp.fit)
```


Panel Model: all candidates, nlp featrues on small donor count
```{r}
dem20.nlp.fit <- plm(small_donor_count_normalized ~ 
  X2018.congress+X2020.democractic.party+X2020.democractic.primary+X2020.democratic.primary+X2020.election+aoc+assange+biden.busing+booker+border+
  busing+buttigieg+civil.rights+climate.change+congress+de.blasio+drugs+economic.foreign.policy+electoral.college+female.congresswomen+first.debate+gun.laws+
  harris+healthcare+hickenlooper+impeachment+inslee+iowa.caucus+iran+iran.afghanistan+israel+kavanaugh+lgbtq+marijuana+mccain+mcconnell+
  mueller.report+north.korea+o.rourke+prison.reform+puerto.rico+russian.interference+sanders+student.debt+supreme.court+tax.return+
  tech+trump+trump.racist+unions+us.economy..socialism.vs.captialism.+venezuela+warren+williamson+yang,
  data=dem20.panel, model="within")

summary(dem20.nlp.fit)
```

Panel Model: all candidates, nlp featrues on donnation amount
```{r}
dem20.nlp.fit <- plm(individual_donation_amount_normalized ~ 
  X2018.congress+X2020.democractic.party+X2020.democractic.primary+X2020.democratic.primary+X2020.election+aoc+assange+biden.busing+booker+border+
  busing+buttigieg+civil.rights+climate.change+congress+de.blasio+drugs+economic.foreign.policy+electoral.college+female.congresswomen+first.debate+gun.laws+
  harris+healthcare+hickenlooper+impeachment+inslee+iowa.caucus+iran+iran.afghanistan+israel+kavanaugh+lgbtq+marijuana+mccain+mcconnell+
  mueller.report+north.korea+o.rourke+prison.reform+puerto.rico+russian.interference+sanders+student.debt+supreme.court+tax.return+
  tech+trump+trump.racist+unions+us.economy..socialism.vs.captialism.+venezuela+warren+williamson+yang,
  data=dem20.panel, model="within")

summary(dem20.nlp.fit)
```

Panel Model: warren, nlp featrues on polls
```{r}
warren <- dem20[dem20$candidate=="warren",]
warren.panel <- pdata.frame(warren, c("date"))
warren.panel <- dem20.panel[!duplicated(rownames(dem20.panel)),]
warren.panel.nlp.fit <- plm(polling_percentage_normalized ~ 
  X2018.congress+X2020.democractic.party+X2020.democractic.primary+X2020.democratic.primary+X2020.election+aoc+assange+biden.busing+booker+border+
  busing+buttigieg+civil.rights+climate.change+congress+de.blasio+drugs+economic.foreign.policy+electoral.college+female.congresswomen+first.debate+gun.laws+
  harris+healthcare+hickenlooper+impeachment+inslee+iowa.caucus+iran+iran.afghanistan+israel+kavanaugh+lgbtq+marijuana+mccain+mcconnell+
  mueller.report+north.korea+o.rourke+prison.reform+puerto.rico+russian.interference+sanders+student.debt+supreme.court+tax.return+
  tech+trump+trump.racist+unions+us.economy..socialism.vs.captialism.+venezuela+warren+williamson+yang,
  data=warren.panel, model="within")

summary(warren.panel.nlp.fit)
```

```{r}
warren_plot <- warren[c("warren","israel","kavanaugh","date","polling_percentage_normalized","aoc")]
#"tv_mentions_normalized"
#"individual_donations_normalized"
warren_plot$israel <- warren_plot$israel * 10
warren_plot = melt(warren_plot, id.vars = 'date')
gg_warren <- ggplot(warren_plot, aes(x=date, y=value)) +
    geom_line(aes(colour=variable)) +
    ggtitle("Warren") +
    xlab("Date") +
    ylab("Y")
gg_warren
```

